A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2017-03-30 casm.tex
some typos, better wording on the concurrent games

The approach we follow here articulates game theory
together with complex network analysis. Putting together game-theoretic modeling
on complex networks is common in studies about the emergence of cooperation
\cite{SzaboFath:2007,Nowak-etal:92,OshtukiNowak:2006,Santos-etal:2005,Santos-etal:2006}.
However, in the context of science studies, it is far from being a unified,
active field of research. There is no common set of questions and problems or a
coherent body of models and theory aimed to understand the quirks and twists of
cooperation in scientific communities. It is more likely to say that work of
this kind are scarce and their implementatoin is on different situations related
to the social life of science, where sometimes models are run on science
collaboration networks because access to these kind of data is relatively easy
to access.\\

@@

Game theory over complex scientific information and collaboration networks has
also attracted attention. Hanauske has applied \emph{Evolutionary Game Theory}
to have a better understanding of the market of scientific information, that
involves scientists as producers but also journal, that are in charge of its
dissemination. One of his models connects scientist interest in maximazing their
reputation and their decissions to publish in an Open Access Journal or in a
journal that works under the traditional publication framework. He has also
modeled a game-theretic interactions between scientists and journals for both
players to go open access or not. Only this second model incorporates networks
analysis \cite{Hanauske-etal:2012}.\\

Another interesting behavior explored by game-theoretical modeling and complex
networks has been done by Li and Cao \cite{Li-etal:2009}. This authors analyze the role of
punishment in a science collaboration network. More precisely, they implemented
an ultimatum bargaining model to understand the impact of the topology of a
social network and the emergence of a population of punishers. In the ultimatum
game, there are two players, one of them (the proponent) suggest to split a sum
in two and its connected neighbor (the respondent), can decide to accept or
reject the offer as it may be perceived more or less fair. If the respondent
accepts, the sum is devided and the shares are divided according to the
offer. If the offer is rejected the players do not get anything. According to
the authors, the act of rejecting a selfish offer can be interpreted as a
punishment since the player that rejects it is paying a cost of not receiving
anything but the proponent looses even more, since his offer was selfish. Their
results show that the population of punishers will emerge as a phase transition
as the degree of selfishness increses among players. More over, they found a
dependence relation between the tolerance to selfish behaviors and the
topologies of networks.\\

@@

   In the context of our model, utility represents access to resources shared by
   others.  The value of the utility function for a player is the sum of the
   payoffs of playing prisoner's dilemma with its neighbors. The other
   concurrent game consists of players trying to maximize the strength of their
   interactions with other players. In the coordination game the best strategy
   is to adopt the same strategy as the other player, as it pays the most,
   regardless of cooperation or defection in the utility game. When both
   cooperate the interaction gets a positive payoff, when both defect, the
   interaction doesn't get affected; but if they do not coordinate, then the
   interaction loses. The payoffs resulting from the coordination games are not
   part of the utility function, however, both are considered by the agent to
   decide whether to cooperate or defect. \\

>>>>

The approach we follow here articulates game theory together with
complex network analysis. Putting together game-theoretic modeling on
complex networks is common in studies about the emergence of
cooperation
\cite{SzaboFath:2007,Nowak-etal:92,OshtukiNowak:2006,Santos-etal:2005,Santos-etal:2006}.
However, in the context of science studies, it is far from being a
unified, active field of research. There is no common set of questions
and problems or a coherent body of models and theory aimed to
understand the quirks and twists of cooperation in scientific
communities. It is more likely to say that work of this kind is
scarce and their implementatoin is on different situations related to
the social life of science, where sometimes models are run on science
collaboration networks because access to these kind of data is
relatively easy to access.\\

@@

Game theory over complex scientific information and collaboration
networks has also attracted attention. Hanauske has applied
\emph{Evolutionary Game Theory} to have a better understanding of the
market of scientific information, that involves scientists as
producers but also journal, that are in charge of its
dissemination. One of his models connects scientist interest in
maximazing their reputation and their decissions to publish in an Open
Access Journal or in a journal that works under the traditional
publication framework. He has also modeled a game-theoretic
interactions between scientists and journals for both players to go
open access or not. Only this second model incorporates networks
analysis \cite{Hanauske-etal:2012}.\\

Another interesting behavior explored by game-theoretical modeling and
complex networks has been done by Li and Cao \cite{Li-etal:2009}. This
authors analyze the role of punishment in a science collaboration
network. More precisely, they implemented an ultimatum bargaining
model to understand the impact of the topology of a social network and
the emergence of a population of punishers. In the ultimatum game,
there are two players, one of them (the proponent) suggest to split a
sum in two and its connected neighbor (the respondent), can decide to
accept or reject the offer as it may be perceived more or less
fair. If the respondent accepts, the sum is divided according to the
offer. If the offer is rejected both players do not get
anything. According to the authors, the act of rejecting a selfish
offer can be interpreted as a punishment since the player that rejects
it is paying a cost of not receiving anything but the proponent looses
even more, since his offer was selfish. Their results show that the
population of punishers will emerge as a phase transition as the
degree of selfishness increses among players. Moreover, they found a
dependence relation between the tolerance to selfish behaviors and the
topologies of networks.\\

@@

In the context of our model, utility represents access to resources
shared by others. The value of the utility function for a player is
the sum of the payoffs of playing prisoner's dilemma with its
neighbors. In the other concurrent game players try to maximize the
strength of their interactions by coordinating their strategy to that
of their neighbors. Interaction strength is updated with no regard to
cooperation or defection in the utility game. When both cooperate the
interaction gets a positive payoff, when both defect, the interaction
doesn't get affected; but if they do not coordinate, then the
interaction loses. The payoffs resulting from the coordination games
are not part of the utility function, however both are considered by
the agent to decide whether to cooperate or defect. \\
